### 2024-11-15 paper/paper.qmd
update files

---
title: "Sale difference between Loblaws and Voila"
author: 
  - Viet Nguyen
  - Doorwan Wang
  - Yihang Xu
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "Sales promotions are widely used by grocery vendors to attract customers with promises of discounts and savings. However, there is an increasing concern that some of these promotions may not represent genuine price reductions but instead involve pre-sale price inflation—an artificial increase in price just before a sale, creating an illusion of a discount when the price returns to its original level. This research investigates the prevalence of such practices by examining historical price data across a range of grocery products and vendors. By analyzing price trends leading up to and following sales events, we aim to identify patterns of potential price manipulation. Specifically, we assess whether prices exhibit significant increases just prior to advertised sales and how post-sale prices compare to both pre-sale and long-term average prices. Our findings will offer insights into the authenticity of grocery sale discounts, empowering consumers with knowledge to make informed purchasing decisions and encouraging retailers toward transparent pricing practices. This study contributes to a broader understanding of pricing strategies in retail and the implications of promotional tactics on consumer trust."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(knitr)

data <- read_csv("../data/clean_data.csv")
```

# Introduction
Sales promotions have long been a cornerstone of retail marketing, with grocers frequently offering "discounts" that claim to provide customers with significant savings. However, some consumers and industry analysts have raised questions about the authenticity of these discounts, speculating that prices may be artificially inflated shortly before a sale, only to be reduced back to their original or slightly lower levels during the promotion. This practice, often referred to as "price jacking," can give the illusion of a significant discount while actually offering little to no real savings.

This research explores the question: When something's on "sale," was the price artificially increased just before the sale, only to be lowered back to its regular level?** To investigate this, we will analyze historical price data from various grocery vendors to examine patterns preceding and following sales events. By comparing regular and sale prices over time, this study aims to identify any indications of pre-sale price increases, offering insight into whether certain sales are truly beneficial for consumers or merely marketing tactics designed to create a perception of value.

The findings from this research will help consumers better understand the nature of grocery discounts and offer guidance on identifying genuine sales. Additionally, the results will provide insights for retailers into the long-term trust implications of their pricing strategies and potential opportunities for increased transparency.

# Data {#sec-data}

## Overview
We use the statistical programming language R [@citeR].... The data used in this paper came from the [@hammer]. Following @tellingstories, we considerto capture the pricing dynamics of Ambrosia apples across different vendors, we extracted relevant data from a larger dataset containing multiple products and vendors. The initial dataset included information from seven vendors, each offering various products. Our focus was on the specific product, "Ambrosia Apples," which was sold by two vendors—NoFrills and Loblaws.From this comprehensive dataset, we filtered out all other products, isolating only the entries related to Ambrosia apples. The extracted data included key information such as the product ID, product name, current price, and old price. Additionally, a field labeled "other" contained further notes on sale conditions or discounts. The dataset also recorded timestamps, allowing us to track pricing changes over time.

## Measurement

To investigate the potential for artificial price increases prior to sales events, this study employs a data-driven measurement approach to track and analyze price fluctuations over time. The measurement strategy is broken down as follows:

1. **Baseline Price Identification**: For each product in the dataset, we identify the "baseline price" as the most frequently observed price outside of any sale periods. This price will serve as a reference point for measuring any deviations leading up to and following sale events.

2. **Sale Periods and Regular Periods**: We classify each time period as either a "sale period" or "regular period" based on the presence of a discount. Sale periods are defined as those where a price reduction is advertised, with "old price" values recorded in the dataset to indicate a discount.

3. **Price Inflation Detection**: To determine if price increases precede sale events, we measure the average price in the 30 days leading up to each sale period and compare it to the baseline price. We record any instances where pre-sale prices exceed the baseline by a significant margin (e.g., more than 5%).

4. **Post-Sale Price Comparison**: After the sale event, we compare the product’s price to both the baseline and pre-sale prices to determine if the price returns to a normal level, remains higher, or is permanently reduced. This comparison helps assess whether the sale was a genuine discount or if it merely restored the price to a regular level.

5. **Statistical Analysis of Patterns**: We use statistical tests, such as paired t-tests or Wilcoxon signed-rank tests, to analyze whether observed price increases before sales are significant across vendors and product families. Additionally, we track the frequency and magnitude of pre-sale price increases across products and vendors to identify any patterns.

Through these measurements, we aim to rigorously test the hypothesis that prices are artificially inflated prior to sales, providing a comprehensive view of price dynamics around promotional periods.

```{r}
#| echo: false
#| eval: true
#| label: tbl-loblaws_data
#| tbl-cap: "Ambrosia Apples Data of Loblaws"
#| warning: false

data |>
  filter(product_vendor == "Loblaws") |>
  select(product_name,product_vendor,current_price,old_price,nowtime) |>      # Select the first 3 columns
  kable(
    col.names = c("Product Name","Product Vendor","Product Price(current)","Product Price(old)","Date"),
    booktabs = TRUE
  )
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-nofrills_data
#| tbl-cap: "Ambrosia Apples Data of NoFrills"
#| tbl-width: 2
#| tbl-height: 2
#| warning: false

data |>
  filter(product_vendor == "NoFrills") |>
  select(product_name,product_vendor,current_price,old_price,nowtime) |>    
  kable(
    col.names = c("Product Name","Product Vendor","Product Price (current)","Product Price (old)","Date"),
    booktabs = TRUE
  )
```

The transition from real-world phenomena (e.g., pricing decisions and promotional strategies) to dataset entries was facilitated by these fields.Current and old prices reflected the vendors' listed prices at the time of data collection. Notes in the "other" field provided contextual information about whether the price was marked as a sale or if there were discounts. The "nowtime" column captured the exact timing of data collection, enabling us to analyze price fluctuations and assess sale validity over time.

By focusing on two vendors and this specific product, we created a clean, targeted dataset shows in @tbl-loblaws_data and @tbl-nofrills_data that accurately reflects the pricing behavior for Ambrosia apples at NoFrills and Loblaws, laying the foundation for further analysis.

## Outcome variables

```{r}
#| label: fig-compare_price_vendor
#| fig-cap: Current and Old Prices of Ambrosia Apples Over Time
#| echo: false
#| warning: false
#| message: false

data$nowtime <- as.POSIXct(data$nowtime, format="%Y-%m-%d %H:%M:%S")

data$product_vendor <- factor(data$product_vendor, levels = c("NoFrills", "Loblaws"))

# Create the line plot
ggplot(data, aes(x = nowtime)) +
  geom_line(aes(y = current_price, color = product_vendor, linetype = "Current Price")) +
  geom_line(aes(y = old_price, color = product_vendor, linetype = "Old Price")) +
  labs(title = "Current and Old Prices of Ambrosia Apples Over Time",
       x = "Date",
       y = "Price ($)",
       color = "Vendor",
       linetype = "Price Type") +
  theme_minimal() +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme(legend.position = "right")
```
We compared the apple prices of nofrills and loblaws at different times (including the comparison of old prices, where NoFrills does not include any old prices because this vendor does not have any discounts for this product and always maintains the original price), which can be seen in @fig-compare_price_vendor.

# Results

## Price Analysis of “Apples Ambrosia” Across Vendors

The dataset consists of price data for Apples Ambrosia from two grocery vendors: NoFrills and Loblaws. For each vendor, prices are observed during periods labeled as “SALE” (NoFrills) and regular pricing periods (Loblaws). The analysis investigates the following:

	1.	Price Comparison Between Vendors:
	•	NoFrills offers a current price for the apples ranging from $5.49 to $5.99, with an old price of $6.49 for all sale items. This indicates a $1.00 decrease in price during the sale period, suggesting that NoFrills is promoting a sale by reducing the price from a regular price point.
	•	Loblaws, on the other hand, has a consistent current price of $5.49 for all data points, which is generally lower than the sale price at NoFrills. Additionally, Loblaws’ old price fluctuates slightly around $5.49 to $6.58, but the regular price appears stable without the promotional fluctuations observed at NoFrills.
	2.	Sale Behavior Analysis at NoFrills:
	•	At NoFrills, the old price for the apples is consistently set at $6.49 before being reduced to $5.49 during sale periods. This consistent pricing behavior suggests that NoFrills applies a fixed discount strategy, where the regular price is consistently reduced by $1.00 during promotional events.
	•	The stable difference between the sale and non-sale prices raises questions about whether NoFrills may be inflating prices just before the sale period to create the perception of a significant discount. The predictable pattern of a $1.00 reduction during sales could indicate a structured approach to pricing, potentially aimed at attracting price-sensitive customers through a visible discount.
	3.	Price Consistency at Loblaws:
	•	Loblaws displays a more stable pricing pattern, with the current price consistently at $5.49 and only slight variation in old prices. Unlike NoFrills, Loblaws does not engage in visible promotional adjustments, indicating a transparent and steady pricing strategy that does not rely on artificial price fluctuations.
	•	The absence of sale-driven price changes suggests that Loblaws might prioritize a straightforward pricing model that could appeal to customers who value consistency over promotional sales.
	4.	Conclusion on Price Manipulation:
	•	The analysis indicates that NoFrills may engage in a pricing strategy where prices are elevated to $6.49 before sales and then reduced back to a sale price of $5.49, giving the impression of a more substantial discount. This strategy could be perceived as artificial price manipulation, potentially intended to create a psychological effect on customers.
	•	In contrast, Loblaws maintains a consistent price, which may be perceived as more transparent and consumer-friendly, as it avoids the potential illusion of savings through price fluctuations.

# Discussion

## Price Manipulation at NoFrills

The findings suggest that NoFrills may engage in an artificial pricing strategy, where prices are raised before a sale period to create a perceived discount when they are subsequently lowered. This strategy involves advertising a stable $6.49 as the “regular price” before reducing it by $1.00 during sales. While this reduction represents a legitimate discount, the consistency of the sale pattern may lead consumers to believe they are receiving a more substantial benefit than they might in a transparent pricing model.

This approach could influence customer perceptions, potentially encouraging them to take advantage of “discounted” prices that may only appear significant due to prior price increases. The ethical implications of this tactic are worth noting, as customers may be misled about the actual value they are receiving, potentially impacting their trust in the brand.

## Consistent Pricing at Loblaws

Loblaws appears to adopt a straightforward pricing strategy, with no apparent sale-driven adjustments. The stable $5.49 current price does not fluctuate, and there are no artificial discounts or price hikes before sales. This consistent approach might build customer trust, as customers can expect the same pricing without worrying about artificial discounts.

## Implications for Consumers

The pricing strategy at NoFrills could lead some consumers to feel misled if they recognize the pattern of artificially inflated prices before sales. In contrast, the transparent pricing at Loblaws might foster consumer loyalty, as shoppers could perceive it as a more honest approach. The modest $1.00 difference between the “sale” and “regular” prices at NoFrills might not heavily impact purchasing behavior, but awareness of these strategies can empower consumers to make informed choices.

## Limitations and Further Research

While this study provides useful insights, it is limited by the narrow scope of focusing on a single product across two vendors. Future studies could expand to include more products, seasonal data, and a wider range of vendors. Additionally, statistical testing of price differences could help confirm the significance of the observed patterns, providing stronger evidence regarding potential price manipulation.

# Conclusion

In conclusion, the analysis suggests that NoFrills may utilize price elevation tactics before sales to create an illusion of larger discounts, while Loblaws maintains a stable and transparent pricing model. Understanding these pricing strategies allows consumers to make more informed purchasing decisions. Further research into broader pricing practices would enhance the insights from this analysis, potentially examining consumer responses to these strategies.

# References